Influence of Connectivity on Dung Beetle Communities

### 2025-06-14 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
gargantuan writing effort p2

Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species. The N = `r how_many_common` species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1). All but four species were captured in every patch type. _Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches. _Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively). All species were within their native ranges.

When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches(GLMM, P < 2e-6, Table 4). Statistical analysis of abundance focused on the six most abundant species. When looking at counts for our most abundant species, results were variable but overall follow a consistent pattern where the most individuals of a given species are more frequent in the matrix then corridor and winged patches with a dip at the rectangle patch (Table 6). One difference to note is that _Aphodius alloblackburneus_ had much higher counts in matrix patches in contrast to other patch types (Figure 6).

@@

This study advances our understanding of the factors shaping dung beetle community composition in temperate regions of the southeastern United States. In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches. Our main findings emphasized: (1) Dominant species were most abundant in forested matrix, but not equally for all species, some were represented proportionally higher in matrix than in open patch habitat. (2) Species richness and diversity were relatively consistent across all treatments, however variation across sampling blocks is evident.

@@

@@

>>>>

Overall, I collected N = `r total_N` dung beetles belonging to N = `r total_spp` species. The N = `r how_many_common` most dominant species comprised of `r round((count_most_common/total_N)*100,1)`% of all captures: \textit{`r most_common_spp[1,2]`} (N = `r most_common_spp[1,1]`), \textit{`r most_common_spp[2,2]`} (N = `r most_common_spp[2,1]`), \textit{`r most_common_spp[3,2]`} (N = `r most_common_spp[3,1]`), \textit{`r most_common_spp[4,2]`} (N = `r most_common_spp[4,1]`), \textit{`r most_common_spp[5,2]`} (N = `r most_common_spp[5,1]`), and \textit{`r most_common_spp[6,2]`} (N = `r most_common_spp[6,1]`; Table 1). All but six species were captured in every patch type. _Onthophagus concinnus_ was only found in the matrix and winged patches, while _Onthophagus striatulus_ was only captured in matrix habitat and rectangular patches. _Geotrupes blackburnii_ and _Onthophagus tuberculifrons_ were the only species restricted to one patch type (winged and matrix, respectively). All species were within their native ranges.

When comparing the total abundance of beetles (all species combined) across patch types, there were no significant difference with one exception: fewer individuals were captured in rectangular patches(GLMM, P < 0.001, Table 4).

Statistical analysis of abundance focused on the six most abundant species. A generalized linear mixed model identified significant effects for both species ID and patch type on dung beetle abundance (Table 7). The baseline abundance (intercept) corresponds to the abundance of _Aphodius alloblackburneus_ in connected patches. Compared to this baseline, _Canthon vigilans_, _Ateuchus lecontei_, and _Phanaeus igneus_ showed significantly higher abundance  (β = 1.05, 0.73, 0.66 respectively; all p < 0.001), while _Onthophagus pennsylvanicus_ had lower abundance (β = –0.80, p < 0.001).

Species-patch interactions indicated indicated species specific responses to treatments. In matrix plots, the positive effect on abundance was reduced for all other species (e.g., cvig × Matrix β = –0.59, p < 0.001). _Ateuchus lecontei_ and _Phanaeus igneus_ showed increased abundance in rectangle patches relative to expectations (β = 0.40 and 0.47, respectively; p < 0.05). Random effects indicated moderate variability among blocks (standard deviation of intercept = 0.51).

@@

This study advances our understanding of the factors shaping dung beetle community composition in temperate regions of the southeastern United States. In addition, the experimental design enables direct comparisons between populations in continuous matrix habitat and those in both isolated and corridor connected patches. Our main findings emphasized: (1) Dominant species were most abundant in forested matrix, but not equally for all species, some were represented proportionally higher in matrix than in open patch habitat. (2) Patch shape and isolation had less of an influence on species richness which was relatively even on both a patch and block level. (3) Species diversity metrics were also relatively even across patch types however varied widely by block. These results suggest that dung beetle species are fully capable of permeating fragmented landscapes and that habitat type and connectivity shape community compositions, but landscape effects of a larger scale are driving changes in biodiversity.

## Abundance

@@

## Richness

@@

Diversity
